A seamless, client-centric programming model for type safe web applications

### 2014-05-02 paper.tex
Update related works intro, add AFAX section.

Other approaches to this problem exist, notably in the form of the
Conductance\ \cite{conductance}, Duetto\ \cite{duetto} and
Sunroof\ \cite{sunroof} projects, all of which implement some synchronous
abstraction over the network between the client and the server, much like our
solution.

\paragraph{Conductance} Conductance is an application server built on
StratifiedJS, a JavaScript language extension which adds a few niceties such as
cooperative multitasking and more concise syntax for many common tasks.
Conductance uses an RPC-based model for client-server communication, much like
our own, but also adds the possibility for the server to independently transmit
data back to the client through the use of shared variables or call back into
the client by way of function objects received via RPC call, as well as the
possibility for both client and server to seamlessly modify variables located
on the opposite end of the network. Conductance is quite new and has no
relevant publications. It is, however, used for several large scale web
applications.

@@

\paragraph{Duetto} Duetto is a C++ compiler targeting the web, written from
the ground up to produce code for both client and server simultaneously. It
utilizes the new attributes mechanism introduced in C++11\ \cite{sepples11} to
designate functions and data to live on either client or server side. Any calls
to a function on the other side of the network and attempts to access remote
data are implicit, requiring no extra annotations or scaffolding at the call
site. Duetto is still a highly experimental project, its first release being
only a few months old, and has not been published in any academic venue.

@@

\paragraph{Sunroof} In contrast to Conductance and Duetto, Sunroof is an
embedded language. Implemented as a Haskell library, it allows the programmer
to use Haskell to write code which is compiled to JavaScript and executed on
the client. The language can best be described as having JavaScript semantics
with Haskell's type system. Communication between client and server is
accomplished through the use of ``downlinks'' and ``uplinks'', allowing for
data to be sent to and from the client respectively.

@@

@@

>>>>

Several other approaches to the seamless client-server interaction exist. In
general, these attacks tend to be of the ``all or nothing'' variety,
introducing new languages or otherwise requiring custom full stack solutions.
In contrast, our solution can be implemented entirely as a library and is
portable to any pair of compilers supporting typed monadic programming.
Moreover, Haste.App has a quite simple and controlled programming model with a
clearly defined controller, which stands in contrast to most related work which
embraces a more flexible but also more complex programming model.

The more notable approaches to the problem are discussed further in this section.

\paragraph{Conductance} Conductance\ \cite{conductance} is an application
server built on StratifiedJS, a JavaScript language extension which adds a few
niceties such as cooperative multitasking and more concise syntax for many
common tasks.
Conductance uses an RPC-based model for client-server communication, much like
our own, but also adds the possibility for the server to independently transmit
data back to the client through the use of shared variables or call back into
the client by way of function objects received via RPC call, as well as the
possibility for both client and server to seamlessly modify variables located
on the opposite end of the network. Conductance is quite new and has no
relevant publications. It is, however, used for several large scale web
applications.

@@

\paragraph{Duetto} Duetto\ \cite{duetto} is a C++ compiler targeting the web,
written from the ground up to produce code for both client and server
simultaneously.
It utilizes the new attributes mechanism introduced in C++11\ \cite{sepples11}
to designate functions and data to live on either client or server side.
Any calls to a function on the other side of the network and attempts to access
remote data are implicit, requiring no extra annotations or scaffolding at the
call site. Duetto is still a highly experimental project, its first release
being only a few months old, and has not been published in any academic venue.

@@

\paragraph{Sunroof} In contrast to Conductance and Duetto,
Sunroof\ \cite{sunroof} is an embedded language. Implemented as a Haskell
library, it allows the programmer to use Haskell to write code which is
compiled to JavaScript and executed on the client. The language can best be
described as having JavaScript semantics with Haskell's type system.
Communication between client and server is accomplished through the use
of ``downlinks'' and ``uplinks'', allowing for data to be sent to and from the
client respectively.

@@

\paragraph{AFAX} AFAX\ \cite{afax}, an F#-based solution, takes an approach
quite similar to ours, using monads to allow client and server side to coexist
in the same program. Unfortunately, using F# as the base of such a solution
raises the issue of side effects. Since any expression in F# may be side
effecting, it is quite possible with AFAX to perform a side effect on the client
and then attempt to perform some action based on this side effect on the server.
To cope with this, AFAX needs to introduce cumbersome extensions to the F# type
system, making AFAX exclusive to Microsoft's F# compiler and operating system,
whereas our solution is portable to any pair of Haskell compilers.

@@

\bibitem[Petricek, Syme (2007)]{afax}
T. Petricek, and Don Syme. "AFAX: Rich client/server web applications in F#." (2007).
